\section*{Références}


\begin{thebibliography}{99}

\bibitem{Stetoscope2025}
Beaufils, T., Boutet, A., et al. (2025).
\textit{STETOSCOPE: underStand TargEting and manipulaTiOnS via Collaborative Private data collection}.
Proceedings of the ACM on IMWUT.

\bibitem{Melicher2016}
Melicher, W., et al. (2016).
\textit{Online Tracking: A 1-Million-Site Measurement and Analysis}.
Proceedings of the 2016 ACM SIGSAC Conference on Computer and Communications Security.

\bibitem{Hannak2014}
Hannak, A., et al. (2014).
\textit{Measuring price discrimination and steering on e-commerce web sites}.
In Proceedings of the 2014 conference on internet measurement conference (pp. 305-318).

\bibitem{Mathur2021}
Mathur, A., et al. (2021).
\textit{What Makes a Dark Pattern... Dark? Design Attributes, Normative Considerations, and Measurement Methods}.
arXiv:2101.04843.

\bibitem{Gray2023}
Gray, C. M., et al. (2023).
\textit{An Ontology of Dark Patterns Knowledge: Foundations, Definitions, and a Pathway for Shared Knowledge-Building}.
arXiv:2309.09640.

\bibitem{Bielova2023}
Bielova, N. (2023).
\textit{Présentation des enjeux informatiques des dark patterns}.
Vidéo Canal-U, VDN.

\bibitem{GDPR2016}
Regulation (EU) 2016/679 (GDPR). (2016).
\textit{General Data Protection Regulation}.
Official Journal of the European Union.

\bibitem{Kirkman2023}
Kirkman, et al. (2023).
\textit{DarkDialogs: Automatic Detection of Dark Patterns in Cookie Banners}.

\bibitem{Yada2022}
Yada, S., et al. (2022).
\textit{Dark patterns in e-commerce: A dataset and its baseline evaluations}.
IEEE BigData.

\bibitem{Chen2023}
Chen, J., et al. (2023).
\textit{UI Guard: Detecting Dark Patterns in Mobile Applications}.
Proceedings of the 36th Annual ACM Symposium on User Interface Software and Technology.

\bibitem{Venugopalan2025}
Venugopalan, R., et al. (2025).
\textit{FP-Inconsistent: Measurement and Analysis of Fingerprint Inconsistencies in Evasive Bot Traffic}.
IMC '25.

\end{thebibliography}
